---
title: "STA130H1 -- Fall 2018"
author: "N. Moon and N. Taback"
subtitle: Week 5 Practice Problems
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
library(tidyverse)
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE) 
```


# Instructions

## What should I bring to tutorial on October 12?

- R output (e.g., plots and explanations) for Question 1 (a)-(e).  You can either bring a hardcopy or bring your laptop with the output.

# Tutorial Grading

Tutorial grades will be assigned according to the following marking scheme.

|                                    | Mark |
|------------------------------------|------|
| Attendance for the entire tutorial | 1    |
| Assigned homework completion^a^    | 1    |
| In-class exercises                 | 4    |
| Total                              | 6    |



These problems are based on the lesson [Joining Data Frames](joiningdataframes.html).


# Practice Problems

The file `heroes_information_exer.csv` contains some information on superheroes and `super_hero_powers_exer.csv` conatins some information on powers of superheroes.

The following questions are based on data in `heroes_information.csv` and `super_hero_powers.csv`.

## Question 1

(a)  Read both data sets `heroes_information.csv` and `super_hero_powers.csv` into R using `read_csv` from the `tidyverse` library.  


If you are using rstudio.cloud then here is the R code.  

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
hero_info <- read_csv("heroes_information_exer.csv")
hero_power <- read_csv("super_hero_powers_exer.csv")
```

If you are using RStudio on your own computer then use this R code.

```{r}

```



How may variables and observations are in each data frame?

(b)  Suggest a key to join the two data frames?

(c) What proprotion of superheroes in `heroes_information` also have data in `super_hero_powers`?  

(d)  What is the number of observations, average, median, standard deviation, and inter-quartile range of `weight` for superheroes for each category of marksmanship? (HINT: use the `group_by()` function then `summarise()`)

(e) Are superheroes with marksmanship thinner compared to those without marksmanship?  Create a visualization to compare the distribution of weight between superheroes that have marksmanship and those that don't have marksmanship.  Which distribution has more variability?


