---
title: "STA130H1 -- Fall 2018"
author: "N. Moon and N. Taback"
subtitle: Week 5 Practice Problems - Solutions
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
library(tidyverse)
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE) 
```


# Instructions

## What should I bring to tutorial on October 12?

- R output (e.g., plots and explanations) for Question 1 (a)-(e).  You can either bring a hardcopy or bring your laptop with the output.

# Tutorial Grading

Tutorial grades will be assigned according to the following marking scheme.

|                                    | Mark |
|------------------------------------|------|
| Attendance for the entire tutorial | 1    |
| Assigned homework completion^a^    | 1    |
| In-class exercises                 | 4    |
| Total                              | 6    |



These problems are based on the lesson [Joining Data Frames](https://ntaback.github.io/UofT_STA130/Fall2018/week5/joiningdataframes.html).


```{r, eval=FALSE, echo=FALSE}
heroes_information <- read_csv("heroes_information.csv")
super_hero_powers <- read_csv("super_hero_powers.csv")

heroes_info1 <- heroes_information %>% 
  select(name, Alignment, Weight, Publisher) %>% 
  filter(Weight != -99) %>%
  group_by(name) %>% 
  mutate(Weight1 = mean(Weight)) %>% 
  distinct(name, .keep_all = T) %>% #remove duplicates
  select(name, Alignment, Weight1, Publisher) %>%
  rename(Weight = Weight1)
  

super_hero_power1 <- super_hero_powers %>% rename(name = hero_names) %>% select(name, Agility, Flight, Marksmanship) 

write_csv(heroes_info1, "heroes_information_exer.csv")
write_csv(super_hero_power1, "super_hero_powers_exer.csv")
```


# Practice Problems

The file `heroes_information_exer.csv` contains some information on superheroes and `super_hero_powers_exer.csv` conatins some information on powers of superheroes.

The following questions are based on data in `heroes_information.csv` and `super_hero_powers.csv`.

## Question 1

(a)  Read both data sets `heroes_information.csv` and `super_hero_powers.csv` into R using `read_csv` from the `tidyverse` library.  Here is the R code.  How may variables and observations are in each data frame?

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
heroinfo_url <- "https://raw.githubusercontent.com/ntaback/UofT_STA130/master/Fall2018/week5/heroes_information_exer.csv"
heropower_url <- "https://raw.githubusercontent.com/ntaback/UofT_STA130/master/Fall2018/week5/super_hero_powers_exer.csv"

hero_info <- read_csv(heroinfo_url)
hero_power <- read_csv(heropower_url)
glimpse(hero_info)
glimpse(hero_power)
```


(b)  Suggest a key to join the two data frames?

Use `name` as the key since it uniquely identifies observations.

(c) What proprotion of superheroes in `heroes_information` also have data in `super_hero_powers`?  

```{r}
inner_join(hero_info, hero_power, by = "name") %>% head()
inner_join(hero_info, hero_power, by = "name") %>% summarise(n = n())
```

The proportion is 460/487 = `r round(460/487, 2)`.

(d)  What is the number of observations, average, median, standard deviation, and inter-quartile range of `weight` for superheroes for each category of marksmanship? (HINT: use the `group_by()` function then `summarise()`)

```{r}
left_join(hero_info, hero_power, by = "name") %>% 
  group_by(Marksmanship) %>% 
  summarise(n = n(), 
            mean_wt = mean(Weight, na.rm = TRUE), 
            sd_wt = sd(Weight, na.rm = TRUE),
            median_wt = median(Weight), 
            iqr_wt = IQR(Weight))
```

(e) Are superheroes with marksmanship thinner compared to those without marksmanship?  Create a visualization to compare the distribution of weight between superheroes that have marksmanship and those that don't have marksmanship.  Which distribution has more variability?

```{r}
left_join(hero_info, hero_power, by = "name") %>% 
  filter(Marksmanship != "NA") %>%
  ggplot(aes(x = Marksmanship, y = Weight)) + geom_boxplot()
```

Superheroes with marksmanship are thinner compared to those without marksmanship.  The variability in weight is greater in those without marksmanship compared to those without.


